---
title: "Microglia Clustering"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
library(cowplot)
library(Seurat)
library(patchwork)
```

## Load microglia object

```{r}
cd11b.integrated <- readRDS("../02_data/intermediate_rds/mg_int.rds") 

DefaultAssay(cd11b.integrated) <- "RNA"
cd11b.integrated@meta.data %>% group_by(Strain, Genotype) %>% summarise(N=n())

cd11b.integrated$Strain <- str_replace_all(cd11b.integrated$Strain, pattern = "B6J", replacement = "B6")
cd11b.integrated$Group <- str_replace_all(cd11b.integrated$Group, pattern = "B6J", replacement = "B6")
cd11b.integrated$Group <- factor(cd11b.integrated$Group, levels = c("B6_WT","B6_APP/PS1","CAST_WT", "CAST_APP/PS1", 
                                                                    "PWK_WT",  "PWK_APP/PS1", "WSB_WT", "WSB_APP/PS1"))
sum_table <- cd11b.integrated@meta.data %>% group_by(seurat_clusters) %>% summarise(N=n(), ave_nCount_RNA=median(nCount_RNA), ave_nFeature_RNA=median(nFeature_RNA), ave_percent.mt=median(percent.mt))
prop.table(table(Idents(cd11b.integrated), cd11b.integrated$Group), margin = 2)
```



```{r}
Idents(cd11b.integrated) <- "seurat_clusters"

path <- "../03_results/04a_microglia_clustering/"
plot_title="mic_cluster"
i=17 #(PCA dim) 
j=0.6 #(resolution)
```

## UMAP plot

### UMAP plot - strain/genotype combined

```{r, fig.width=4, fig.height=4}
DimPlot(cd11b.integrated, reduction = "umap", label = TRUE, pt.size = 0.001, label.size = 5) + 
  coord_fixed()+
  theme(axis.title = element_blank(), legend.position = "none")
ggsave(paste(path, plot_title, "_", "umap_", i, "_res_", j, "_DimPlot_Strain_small",  ".png", sep=""), units = "in", width = 4, height = 4,  dpi=300)
```


## check marker genes of each cluster, list top 10 marker genes

```{r}

load("../03_results/01c_integration_mg/pca_res_search/int_mg_pca_17_res_0.6_markers_proptbl.rda")
cd11b.markers %>% group_by(cluster) %>%  top_n(-10, wt=p_val_adj)

```

## combine homeostatic microglia

```{r}

cd11b.integrated$final_clusters <-  ifelse(cd11b.integrated$seurat_clusters %in% 0:5,"H",
                                           cd11b.integrated$seurat_clusters %>% as.character())
cd11b.integrated$final_clusters <- factor(cd11b.integrated$final_clusters, 
                                          levels = c("H", "6", "7", "8", "9", "10", "11", "12"))
Idents(cd11b.integrated) <- "final_clusters"
```

```{r, eval=FALSE}
cd11b.markers_new <- FindAllMarkers(cd11b.integrated, only.pos = FALSE, min.pct = 0.2, logfc.threshold = 0.25, max.cells.per.ident=500) #max.cells.per.ident
saveRDS(cd11b.markers_new, file = paste(path, "mg_markers_new.rds", sep=""))

# get rid of ribo genes in marker genes
ribo.genes <- grep(pattern = "^Rp[sl][[:digit:]]", x = rownames(cd11b.integrated@assays$RNA), value = TRUE)
cd11b.markers <- filter(cd11b.markers, !gene %in% ribo.genes)

```

```{r}
cd11b.markers_new <- readRDS(paste(path, "mg_markers_new.rds", sep=""))
cd11b.markers_new <- cd11b.markers_new %>% mutate(gene=paste(" ", gene, sep = ""))
write_delim(cd11b.markers_new, paste(path, "mg_markers_new.txt", sep = ""), delim = "\t")
top_20 <- cd11b.markers_new %>% group_by(cluster) %>% filter(avg_logFC>0.3) %>% top_n(20, wt = avg_logFC)
top_20
### Marks for DAM-1, DAM-2
### vilin plot: Cst7, Lpl, Apoe, Clec7a, Cx3cr1
### FeaturePlot: Cst7

### Marks for IRM
### vilin plot: Ifit3, Ifitm3, Irf7
### FeaturePlot: Ifit3

### Marks for Hexb-high cells
### vilin plot: Hexb, Cd81, Cst3
### FeaturePlot: Hexb


### Marks for Ribosomal-gene-enriched cells
### vilin plot: Rplp1, Rps21, Rps24
### FeaturePlot: Rplp1

### Marks for Ccl3/Ccl4+ cells
### vilin plot: Ccl3, Ccl4, C3ar1
### FeaturePlot: Ccl3


### Marks for proliferative microglia
### vilin plot: Stmn1, Top2a, Birc5
### FeaturePlot: Stmn1
```
### UMAP plot - strain/genotype splitted

```{r, fig.width=6, fig.height=7}
Idents(cd11b.integrated) <- "final_clusters"

p <- DimPlot(cd11b.integrated, reduction = "umap", label = TRUE, pt.size = 1E-10, split.by = "Group", ncol = 2, label.size = 3, repel = TRUE) + 
  coord_fixed()+
  theme(text = element_blank(), legend.position = "none")
ggsave(paste(path, plot_title, "_", "umap_", i, "_res_", j, "_DimPlot_Strain_split_no_text",  ".png", sep=""), p, units = "in", width = 3, height = 5.7,  dpi=300)

DimPlot(cd11b.integrated, reduction = "umap", label = TRUE, pt.size = 1E-10, split.by = "Group", ncol = 2, label.size = 3, repel = TRUE) + 
  coord_fixed()+
  theme(title = element_text(size = 3), 
        axis.text = element_text(size = 6),
        legend.position = "none")
```

## Dot plot of microglia marker genes for each clusters

```{r, fig.width= 6, fig.height=6}

Idents(cd11b.integrated) <- "final_clusters"

genes <- c("Cst7", "Lpl", "Apoe", "Clec7a", "Cx3cr1", "Tmem119", "Ifit3", "Ifitm3", "Irf7", "Hexb", "Cd81", "Cst3", "Rplp1", "Rps21", "Rps24", "Ccl4", "Ccl3", "C3ar1", "Stmn1", "Top2a", "Birc5")

file_name <- paste(path, "Dotplot_all_gene_H_combine.png", sep = "")
DotPlot(cd11b.integrated, features = genes) + RotatedAxis() + 
  theme(axis.title = element_blank()) + 
  scale_y_discrete(labels = function(x) str_wrap(x, width = 20)) + 
  coord_flip() + 
  theme(axis.text.y = element_text(face = "bold.italic"))
ggsave(file_name ,units = "in", width = 5.5, height = 5.5, dpi=300)

```

## Feature plots of marker genes of microglia subclusters
```{r, fig.width=10, fig.height=7}

genes <- c("Hexb", "Cst7", "Ifitm3", "Rplp1", "Ccl4", "Stmn1")

# # plot one gene
# FeaturePlot(cd11b.integrated, features = c("Tmem119"), min.cutoff = "q9",label=TRUE, repel = FALSE, ncol = 2, order = TRUE) + 
#   coord_fixed() +
#   theme(axis.line = element_blank(), 
#         axis.title = element_blank(),
#         axis.text = element_blank(),
#         axis.ticks = element_blank())

p <- genes %>% 
  map(~FeaturePlot(cd11b.integrated, features = ., min.cutoff = "q9",label=TRUE, repel = FALSE, ncol = 2, order = TRUE) + 
        coord_fixed() +
        theme(axis.line = element_blank(), 
              axis.title = element_blank(),
              axis.text = element_blank(),
              axis.ticks = element_blank())
  )

(p[[1]]+p[[2]]+p[[3]])/(p[[4]]+p[[5]]+p[[6]])
ggsave(paste(path, "Feature_plot_all.png", sep=""), units = "in", width = 10, height = 7,  dpi=300)


```

## Check immediate early  genes (IEG)

#### To verify that our microglia prepared by mechniacal dissociation at 4 °C were not globally ex vivo activated

### Feature plots for IEG
```{r}

genes <- c("Fos", "Fosb", "Dusp1", "Nr4a1", "Arc", "Egr1")

# FeaturePlot(cd11b.integrated, features = c("Egr1"), split.by = "Genotype", min.cutoff = "q9",label=TRUE, repel = TRUE, ncol = 2, order=TRUE) # + coord_fixed()

p <- genes %>% 
  map(~FeaturePlot(cd11b.integrated, features = ., min.cutoff = "q9",label=TRUE, repel = TRUE, ncol = 2, order=FALSE) + 
        coord_fixed()+
        theme(axis.line = element_blank(), 
              axis.title = element_blank(),
              axis.text = element_blank(),
              axis.ticks = element_blank()
        )
  )

(p[[1]]+p[[2]]+p[[3]])/(p[[4]]+p[[5]]+p[[6]])
ggsave(paste(path, "Feature_plot_all_unordered.png", sep=""), units = "in", width = 10, height = 7,  dpi=300)

```

### Dot plot for IEG
```{r, fig.width=6.1, fig.height=3.4}
DotPlot(cd11b.integrated, features = genes) + RotatedAxis() + 
  theme(axis.title = element_blank(),
        axis.text.y = element_text(face = "bold.italic")) + 
  scale_y_discrete(labels = function(x) str_wrap(x, width = 20)) + 
  coord_flip() 
ggsave(paste(path, "Dotplot_IEG.png", sep = ""),units = "in", width = 6.1, height = 3.4, dpi=300)

```

## Deeper dive into homeostatic microglia subclusters

### Dot plot showing other activated microlgia marker genes (cluster 6-12) in homeostatic microglia subclusters are not quite different

```{r, fig.width=6.1, fig.height=5.5}
Idents(cd11b.integrated) <- "seurat_clusters"

genes <- c("Cst7", "Lpl", "Apoe", "Clec7a", "Cx3cr1", "Tmem119", "Ifit3", "Ifitm3", "Irf7", "Hexb", "Cd81", "Cst3", "Rplp1", "Rps21", "Rps24", "Ccl4", "Ccl3", "C3ar1", "Stmn1", "Top2a", "Birc5")

DotPlot(cd11b.integrated %>% subset(idents = c(0:5, 8)), features = genes) + RotatedAxis() + 
  theme(axis.title = element_blank(),
        axis.text.y = element_text(face = "bold.italic")) + 
  scale_y_discrete(labels = function(x) str_wrap(x, width = 20)) + 
  coord_flip() 
ggsave(paste(path, "Dotplot_all_gene_homeo.png", sep = "") ,units = "in", width = 6.1, height = 5.5, dpi=300)
paste(path, "Dotplot_all_gene_homeo.png", sep = "")

### Marks for homeostatic sub microlgia: 
#0 Rpl15-ps2
#1 P2ry12, Sfi1, Rps18-ps3
#2 Malat1, Macf1, Xist, Abca9, Tmem72
#3 Rpl15-ps3, Crybb1, Rpl37rt
#4 Rpl12, Rpl23a
#5 Ifngr1, Txnip, Glul
#8 C1qa, C1qb, Lgmn
```

### Dot plot showing top marker genes in homeostatic microglia subclusters are mostly ribosomal genes or pseudogenes

```{r, fig.width=6.1, fig.height=5.5}
genes_homeo <- c("Rpl15-ps2","P2ry12", "Sfi1", "Rps18-ps3", "Malat1", "Macf1", "Xist", "Abca9", "Tmem72",
           "Rpl15-ps3", "Crybb1", "Rpl37rt", "Rpl12", "Rpl23a", "Ifngr1", "Txnip", "Glul", 
           "C1qa", "C1qb", "Lgmn", "Itm2b")
                                                        
DotPlot(cd11b.integrated %>% subset(idents = c(0:5, 8)), features = genes_homeo) + RotatedAxis() + 
  theme(axis.title = element_blank(),
        axis.text.y = element_text(face = "bold.italic")) + 
  scale_y_discrete(labels = function(x) str_wrap(x, width = 20)) + 
  coord_flip() 
ggsave(paste(path, "Dotplot_all_gene_homeo2.png") ,units = "in", width = 6.1, height = 5.5, dpi=300)
```

## Check the gene number, percent of ribosomal genes and percent of mitochodrial genes in each cluster

```{r, fig.width=4.5, fig.height=4}

QC_plot_single2 <-function(data, y){
  p <- ggplot(data, aes_string(x="final_clusters", y=y, color= "final_clusters")) +
    geom_violin() +
    geom_boxplot(width=0.07, outlier.shape = NA, color = "black", alpha=0.7) +
    theme_bw()+
    theme(legend.position = "none", axis.title.x = element_blank(), axis.ticks = element_blank())
  return(p)
}

p_QC <- c("nFeature_RNA", "percent.mt", "percent.ribo") %>% map(~QC_plot_single2(cd11b.integrated@meta.data, .))

p <- plot_grid(plotlist=p_QC, nrow=3, ncol = 1, align="hv")
plot_grid(p, nrow=1, rel_heights=c(0.1, 1))
ggsave(paste(path, "all_microglia_integrated2.png", sep=""), units = "in", width = 4.5, height = 3.5,  dpi=300)

```




