---
title: "Microglia  Statistics"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
library(cowplot)
library(Seurat)

path = "../03_results/04b_microglia_statistics/"
```

### Load object
```{r}
cd11b.integrated <- readRDS("../02_data/intermediate_rds/mg_int.rds")  # i=17 (PCA dim) # j=0.6 (resolution)

DefaultAssay(cd11b.integrated) <- "RNA"
cd11b.integrated@meta.data %>% group_by(Strain, Genotype) %>% summarise(N=n())

cd11b.integrated$Strain <- str_replace_all(cd11b.integrated$Strain, pattern = "B6J", replacement = "B6")
cd11b.integrated$Group <- str_replace_all(cd11b.integrated$Group, pattern = "B6J", replacement = "B6")
cd11b.integrated$Group <- factor(cd11b.integrated$Group, levels = c("B6_WT","B6_APP/PS1","CAST_WT", "CAST_APP/PS1",
                                                                    "PWK_WT",  "PWK_APP/PS1", "WSB_WT", "WSB_APP/PS1"))
cd11b.integrated$batch <- factor(cd11b.integrated$batch, levels = c("D", "C", "B", "A"))


sum_table <- cd11b.integrated@meta.data %>% group_by(seurat_clusters) %>% summarise(N=n(), ave_nCount_RNA=median(nCount_RNA), ave_nFeature_RNA=median(nFeature_RNA), ave_percent.mt=median(percent.mt))
prop.table(table(Idents(cd11b.integrated), cd11b.integrated$Group), margin = 2)
```

### plot both genotypes in all strains (all replicates combined)
```{r, fig.width=4.5, fig.height=4.5}
# generate meta data, combine cluster 0-5 into H for fraction plot (all replicates combined)
cd11b.meta <- cd11b.integrated@meta.data %>% 
  mutate(Genotype=factor(Genotype, levels=c("WT", "APP/PS1")),
         Strain=factor(Strain, levels = c("B6", "CAST", "PWK", "WSB")),
         new_clusters=ifelse(seurat_clusters %in% 0:5, "H", as.character(seurat_clusters)),
         new_clusters=factor(new_clusters, levels = c("H", "6", "7", "8", "9", "10", "11", "12"))) %>% 
  group_by(Strain, Genotype,new_clusters) %>% 
  summarise(N=n())

ggplot(cd11b.meta, aes(y=N, x=Genotype, fill= new_clusters)) + 
  geom_bar(stat="identity", position="fill", color="black") + 
  labs(y="Fraction", fill = "Clusters") +
  facet_grid(~ Strain) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = c("bold", "bold.italic")),
        axis.title.x = element_blank(), 
        strip.text.x = element_text(face = "bold"), 
        axis.ticks.x = element_blank(), 
        axis.line.x = element_blank())
ggsave(paste(path, "fraction_WT_APPPS1.png", sep=""), width = 4.5, height = 4.5, units = "in")

# Display fraction for H combined clusters
cd11b.integrated$new_clusters <- ifelse(cd11b.integrated$seurat_clusters %in% 0:5, "H", as.character(cd11b.integrated$seurat_clusters))
Idents(cd11b.integrated) <- "new_clusters"
x <- prop.table(table(Idents(cd11b.integrated), cd11b.integrated$Group), margin = 2)
x

# Display the total number of N for each clusters in each group
y <- cd11b.meta %>% filter(new_clusters %in% c("H", "6", '7', '12')) %>% arrange(new_clusters)
y
```

### plot both genotypes in all strains (all replicates separated)
```{r, fig.width=3.5, fig.height=5}
### generate meta data, combine cluster 0-5 into H for fraction plot (all replicates separated)
cd11b.meta <- cd11b.integrated@meta.data %>% 
  mutate(Genotype=factor(Genotype, levels=c("WT", "APP/PS1")),
         Strain=factor(Strain, levels = c("B6", "CAST", "PWK", "WSB")),
         new_clusters=ifelse(seurat_clusters %in% 0:5, "H", as.character(seurat_clusters)),
         new_clusters=factor(new_clusters, levels = c("H", "6", "7", "8", "9", "10", "11", "12"))) %>% 
  group_by(rep, Strain, Genotype, Group, batch, new_clusters) %>% 
  arrange(Group) %>%
  summarise(N=n())

p <- ggplot(cd11b.meta, aes(y=N, x=batch, fill= new_clusters)) + 
  geom_bar(stat="identity", position=position_fill(), color="black") + 
  labs(y="Fraction", fill = "Clusters") +
  facet_grid(Group~., switch="y")+
  coord_flip()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(), 
        strip.text = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.line.y = element_blank(), 
        legend.position = "none")
ggsave(paste(path, "fraction_replicates_seperated.png", sep=""), p, width = 3.5, height = 5, units = "in")

# top to bottome: "B6_WT","B6_APP/PS1","CAST_WT", "CAST_APP/PS1", "PWK_WT",  "PWK_APP/PS1", "WSB_WT", "WSB_APP/PS1"
```

### Box plot for all microglia
#### generate meta data, combine cluster 0-5 into H for statistical testing and box plot
```{r}
cd11b.meta.stat <- cd11b.integrated@meta.data %>% 
  mutate(Genotype=factor(Genotype, levels=c("WT", "APP/PS1")),
         Strain=factor(Strain, levels = c("B6", "CAST", "PWK", "WSB")),
         new_clusters=ifelse(seurat_clusters %in% 0:5, "H", as.character(seurat_clusters)),
         new_clusters=factor(new_clusters, levels = c("H", "6", "7", "8", "9", "10", "11", "12"))) %>% 
  group_by(Strain, Genotype,rep, new_clusters) %>% 
  summarise(Med_nFeature=median(nFeature_RNA), 
            Med_percent_mt=median(percent.mt), 
            Med_percent_ribo=median(percent.ribo),
            N=n()) %>% 
  group_by(rep,Percent=N/sum(N)*100)

```


```{r, fig.width=4, fig.height=10}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

cd11b.meta.stat %>%
  ggplot(aes(y=Percent, x=Genotype, color=Genotype)) +
  geom_boxplot(outlier.size = 0, alpha=0.5) +
  geom_point(aes(color=Genotype), position=position_jitterdodge(), alpha=0.8) + 
  scale_colour_manual(values=cbPalette) + 
  theme_bw() +
  facet_grid(new_clusters ~ Strain, scales= "free_y")+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        legend.text = element_text(size = 10),
        strip.text = element_text(face = "bold", size = 12), 
        axis.ticks.x = element_blank(), 
        axis.title= element_blank(), 
        #legend.text = element_text(face = c("plain", "italic")),
        legend.position = "bottom")
ggsave(paste(path, "cluster_box_all.png", sep=""), width = 4, height = 10, units = "in")

```

### Perform two-way ANOVA to determine the effect of strain and genotype on the percent of micrglia subclusters 
```{r}
####single function for stain and GT interaction debugging

clusters <- unique(cd11b.meta.stat$new_clusters) %>% as.list()

data = cd11b.meta.stat %>% filter(new_clusters %in% clusters[[1]])
aov_object = aov(Percent ~ Strain*Genotype , data=data)
aov.pvals = summary(aov_object)
aov.pvals= aov.pvals[[1]][5] %>% t() %>% as.data.frame()
names(aov.pvals) <- c("Strain", "Genotype", "Strain_Genotype", "Residuals")
aov.pvals <- aov.pvals %>% 
  select(-Residuals) %>% 
  mutate(Cluster = clusters[1] %>% as.character())

aov_StrainGT <- function(cluster, data){
  data = data %>% filter(new_clusters %in% cluster)
  aov_object = aov(Percent ~ Strain*Genotype, data=data)
  aov.pvals = summary(aov_object)
  aov.pvals= aov.pvals[[1]][5] %>% t() %>% as.data.frame()
  names(aov.pvals) <- c("Strain", "Genotype", "Strain_Genotype", "Residuals")
  aov.pvals <- aov.pvals %>% 
    select(-Residuals) %>% 
    mutate(Cluster = cluster %>% as.character())
  return(aov.pvals)
}

aov_StrainGT_object <- function(cluster, data){
  data = data %>% filter(new_clusters %in% cluster)
  aov_object = aov(Percent ~ Strain*Genotype, data=data)
  return(aov_object)
}

aov_StrainGT_table <- clusters %>% map_df(aov_StrainGT, data=cd11b.meta.stat)
aov_StrainGT_table <- aov_StrainGT_table %>% mutate_if(is.double, p.adjust)

aov_StrainGT_table$Cluster[aov_StrainGT_table$Strain_Genotype<0.05]
# [1] "H"  "6"  "7"  "12"

aov_StrainGT_table$Cluster[aov_StrainGT_table$Strain<0.05]
# [1] "6"  "7"  "9" "11" "12" 

aov_StrainGT_table$Cluster[aov_StrainGT_table$Genotype<0.05]
# [1] "6"  "7"  "H" "10" "11" "12"

# keep the annova object for  
aov_object_list <- clusters %>% map(aov_StrainGT_object, data=cd11b.meta.stat)
names(aov_object_list) <- clusters %>% unlist()
TukeyHSD(aov_object_list[["H"]]) %>% .$`Strain:Genotype` %>% data.frame(.,cluster="H")

# to export the statistic result: 
stat_list <- vector(mode = "list", length = length(clusters %>% unlist()))
names(stat_list) <- clusters %>% unlist()
for (i in clusters %>% unlist()){
  stat_list[[i]] <-  TukeyHSD(aov_object_list[[i]]) %>% 
    .$`Strain:Genotype` %>% 
    data.frame(.,cluster=i) %>% 
    rownames_to_column(var = "comparison")
}

stat_all <- do.call(rbind, stat_list)

# find genetype difference within strain comparison: 
stat_APP_cluster <- stat_all %>% 
  filter(comparison %in% c("B6:APP/PS1-B6:WT", "CAST:APP/PS1-CAST:WT", "PWK:APP/PS1-PWK:WT", "WSB:APP/PS1-WSB:WT")) %>% 
  mutate(Significance=ifelse(p.adj<0.05, "S", "NS"))
write_delim(stat_APP_cluster, paste(path, "stat_GT_within_strain.txt", sep = ""), delim = "\t")

# find strain difference of WT (comparing to B6 in for WT in each strain)
stat_WT_strain_cluster <- stat_all %>% 
  filter(str_count(comparison, "WT")==2) %>% 
  mutate(Significance=ifelse(p.adj<0.05, "S", "NS"))
write_delim(stat_WT_strain_cluster, paste(path, "stat_WT_between_strain.txt", sep = ""), delim = "\t") 

# find strain difference of APP/PS1 (comparing to B6 in for APP/PS1 in each strain)
stat_APP_strain_cluster <- stat_all %>% 
  filter(str_count(comparison, "APP/PS1")==2) %>% 
  mutate(Significance=ifelse(p.adj<0.05, "S", "NS"))
write_delim(stat_APP_strain_cluster, paste(path, "stat_APP_between_strain.txt", sep = ""), delim = "\t") 

```


### check the statistics on nFeature, percent of mitochodria, and percent of ribosomal genes for each cluster
```{r}

## nFeature
aov_stat = aov(Med_nFeature ~ new_clusters, data=cd11b.meta.stat)
aov_table <- TukeyHSD(aov_stat) %>% .$new_clusters %>% data.frame() %>% rownames_to_column(var = "comparison") %>% 
  mutate(comparison=paste(" ", comparison, sep = ""), Significance=ifelse(p.adj<0.05, "S", "NS"))
write_delim(aov_table, path = paste(path, "Med_nFeature_comp_cluster.txt", sep = ""), delim = "\t")
filter(aov_table, Significance=="S")

## percent of mitochodria
aov_stat = aov(Med_percent_mt ~ new_clusters, data=cd11b.meta.stat)
aov_table <- TukeyHSD(aov_stat) %>% .$new_clusters %>% data.frame() %>% rownames_to_column(var = "comparison") %>% 
  mutate(comparison=paste(" ", comparison, sep = ""),Significance=ifelse(p.adj<0.05, "S", "NS"))
write_delim(aov_table, path = paste(path, "Med_percent_mt_comp_cluster.txt", sep = ""), delim = "\t")
filter(aov_table, Significance=="S")

## percent of ribosomal
aov_stat = aov(Med_percent_ribo ~ new_clusters, data=cd11b.meta.stat)
aov_table <- TukeyHSD(aov_stat) %>% .$new_clusters %>% data.frame() %>% rownames_to_column(var = "comparison") %>% 
  mutate(comparison=paste(" ", comparison, sep = ""),Significance=ifelse(p.adj<0.05, "S", "NS"))
write_delim(aov_table, path = paste(path, "Med_percent_ribo_comp_cluster.txt", sep = ""), delim = "\t")
filter(aov_table, Significance=="S")


```

## Pseudotime analysis (diffusion map)
#### too many cells for diffusion map, need sampling

```{r, fig.width=3.5, fig.height=5.3}
library(destiny)
library(SingleCellExperiment)
library(scater)

cd11b.integrated$final_clusters <-  ifelse(cd11b.integrated$seurat_clusters %in% 0:5,"H",
                                           cd11b.integrated$seurat_clusters %>% as.character())

sampling <- cd11b.integrated@meta.data %>% 
  rownames_to_column(var = "cell_ID") %>% 
  group_by(Group) %>% 
  sample_n(1000) # take 1000 random cells from each group

mg.small <- subset(cd11b.integrated, cells=sampling$cell_ID)

mg.small <- as.SingleCellExperiment(mg.small)

# use diffusion map to calculate pseudotime
pca <- reducedDim(mg.small)
cellLabels <- mg.small$seurat_clusters

pca_tidy <- as.data.frame(pca) %>% rownames_to_column()

rownames(pca) <- cellLabels

dm <- DiffusionMap(pca)

dpt <- DPT(dm) #

mg.small$pseudotime_dpt <- rank(dpt$dpt) 

df <- colData(mg.small) %>% as.data.frame()

df$final_clusters<- ifelse(df$seurat_clusters %in% 0:5,"H", df$seurat_clusters %>% as.character())

ggplot(df, aes(pseudotime_dpt, fill=final_clusters)) +
  geom_histogram(binwidth = 100, color="grey",size=0.1)+ 
  facet_grid(Group~., switch="y")+ 
  scale_y_continuous("count", position="right") +
  labs(x="DAM <-  pseudotime  -> Homeostatic")+
  theme_bw()+
  theme(text = element_text(family = "Arial", size = 10),
        strip.text.y = element_text(size=5),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y.right = element_blank(),
        legend.position = "null")
ggsave(paste(path, "pseudotime.png", sep = ""), width = 3.5 , height = 5.3, units = "in", dpi = 600)

```




