---
title: "Microglia annotation (AddModule methods)"
output: 
  html_document:
    keep_md: true
---



```{r global_options}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## Compare microglia from other datasets with ours using AddModuleScore function in Seurat
#### Beth Steven's aging dataset
#### Bart de Strooper's AD dataset
#### Ido Amit's AD dataset

```{r} 
library(Seurat)
library(patchwork)
library(tidyverse)
```



### Load our dataset

```{r}
output_path <- "../03_results/07a_mic_compare_ms_annotation/"

cd11b.integrated <- readRDS("../02_data/intermediate_rds/mg_int.rds") 

## combine homeostatic microglia
cd11b.integrated$final_clusters <-  ifelse(cd11b.integrated$seurat_clusters %in% 0:5,"H",
                                           cd11b.integrated$seurat_clusters %>% as.character())
cd11b.integrated$final_clusters <- factor(cd11b.integrated$final_clusters, 
                                          levels = c("H", "6", "7", "8", "9", "10", "11", "12"))
cd11b.integrated$Genotype <- factor(cd11b.integrated$Genotype, levels = c("WT", "APP/PS1"))

cd11b.integrated$Strain <- gsub("B6J", "B6", cd11b.integrated$Strain)
cd11b.integrated$Strain <- factor(cd11b.integrated$Strain, levels = c("B6", "CAST", "PWK", "WSB"))

cd11b.integrated$Group <- gsub("B6J", "B6", cd11b.integrated$Group)

cd11b.integrated$Group <- factor(cd11b.integrated$Group, 
                                 levels=c("B6_WT", "B6_APP/PS1", "CAST_WT", "CAST_APP/PS1", "PWK_WT","PWK_APP/PS1", "WSB_WT", "WSB_APP/PS1"))

```


### Load marker genes of DAM in Ido Amits dataset
select top DAM marker genes

```{r}
df_comp <- readxl::read_xlsx("../02_data/mic_comp_Amit/Amit_HvsDAM.xlsx", col_types=c("text", rep("numeric",4)))

## The spread sheet is from Table S3 of Ido Amit's DAM paper: https://www.cell.com/fulltext/S0092-8674(17)30578-0
## Original link of the dataset is:
# "https://www.cell.com/cms/10.1016/j.cell.2017.05.018/attachment/84d2cdbe-568b-456d-8ab5-a5580ac075b6/mmc4.xlsx"

names(df_comp) <- c("Symbol", "UMI_H", "log2FC", "log10p", "log10FDR")

df_comp <- mutate(df_comp, 
                      p_val=10^(-log10p), 
                      p_val_adj=10^(-log10FDR))

df_comp$log2FC %>% hist() 
df_comp$log2FC %>% summary()

df_comp$log10FDR %>% hist()
df_comp$log10FDR %>% summary()


## DAM markers
top <- 50
df_DAM <- df_comp %>% filter(log10FDR>10, log2FC>0.5, str_detect(Symbol, "^Gm", negate = TRUE)) %>% arrange(desc(log10FDR)) %>% head(50) ## top 49 genes

df_DAM
gene_DAM <- df_DAM$Symbol %>% unlist()

gene_Homeo <- c("Tmem119", "Cx3cr1", "P2ry12", "P2ry13", "Cx3cr1", "Olfml3",  "Selplg", "Siglech", "Csf1r", "Cst3", "Sparc", "C1qa", "C1qb", "C1qc", "Tmsb4x", "Sall1", "Fcrls", "Gpr34", "Spi1",  "Mafb", "Maf", "Mef2a", "Irf8") 
```


```{r}
cd11b.integrated <- AddModuleScore(object = cd11b.integrated, features = list(gene_DAM), ctrl = 100, name = 'DAM_feature') 

cd11b.integrated <- AddModuleScore(object = cd11b.integrated, features = list(gene_Homeo), ctrl = 100, name = 'Homeostatic_feature') 

```


### Load marker genes of ARM, IRM, CPM in Strooper's dataset
#### The marker genes is not readily uploaded onto the website, we recovered their dataset and calculated marker genes

```{r}
df_comp <- read_delim("../03_results/06a_mic_Strooper/Seurat_Strooper_markers.txt", delim = "\t")

df_comp %>% group_by(cluster) %>% top_n(n=-10, wt=p_val_adj)

```

### ARM, IRM, CPM
```{r}

gene_ARM <- df_comp %>% filter(cluster==2, -log10(p_val_adj)>10, avg_logFC>0.5, str_detect(gene, "^Gm", negate = TRUE)) %>% arrange(desc(-log10(p_val_adj))) %>% .$gene %>% unlist() %>% head(top) 
gene_IRM <- df_comp %>% filter(cluster==7, -log10(p_val_adj)>10, avg_logFC>0.5, str_detect(gene, "^Gm", negate = TRUE)) %>% .$gene %>% unlist() %>% head(top)
gene_CPM <- df_comp %>% filter(cluster==6, -log10(p_val_adj)>10, avg_logFC>0.5) %>% .$gene %>% unlist() %>% head(top)
gene_CPM <- c(gene_CPM, "Top2a", "Mcm2") # 4 genes adding original data.

```

```{r}
feature="ARM_feature"
cd11b.integrated <- AddModuleScore(object = cd11b.integrated, features = list(gene_ARM), ctrl = 100, name = feature) 

feature="IRM_feature"
cd11b.integrated <- AddModuleScore(object = cd11b.integrated, features = list(gene_IRM), ctrl = 100, name = feature) 

feature="CPM_feature"
cd11b.integrated <- AddModuleScore(object = cd11b.integrated, features = list(gene_CPM), ctrl = 100, name = feature) 

```



### Load marker genes of OA2 (Ccl3/Ccl4 enriched), OA3 (IRM equivalent) in Steven's dataset
```{r}

input_file <- "../02_data/mic_comp_Stevens/Stevens_cluster_marker.xlsx"
M_stevens_sheet_name <- readxl::excel_sheets(input_file)

sheet <- 1:(length(M_stevens_sheet_name)-1) # do not include the last Transcription factors

df_comp <- vector(mode = "list", length = length(sheet))

for(i in sheet){
  df_comp[[i]] <- readxl::read_xlsx(input_file, sheet = i, skip = 2)
  df_comp[[i]] <- df_comp[[i]] %>% 
    select(1:6)%>% 
    mutate(cluster=M_stevens_sheet_name[i]) 
}

df_comp <- do.call(rbind, df_comp)

colnames(df_comp) <- c("gene", "FC", "p_val", "p_val_adj", "pct_1", "pct_2", "cluster")

df_comp <- df_comp %>% mutate(cluster=str_remove_all(cluster, "Cluster|\\(|\\)| "), 
                              p_val=as.double(p_val),
                              p_val_adj=as.double(p_val_adj))
df_comp$cluster %>% unique()
```

```{r}
gene_OA2 <- df_comp %>% filter(cluster=="AgingOA2", -log10(p_val_adj)>10, log2(FC)>0.5, str_detect(gene, "^Gm", negate = TRUE)) %>% arrange(p_val_adj, desc(FC)) %>% head(50) %>% .$gene %>% unlist() # 50 genes

gene_OA3 <- df_comp %>% filter(cluster=="AgingOA3", -log10(p_val_adj)>10, log2(FC)>0.5, str_detect(gene, "^Gm", negate = TRUE)) %>% arrange(p_val_adj, desc(FC)) %>% head(50) %>% .$gene %>% unlist() # 19 genes


feature="Aging_OA2"
cd11b.integrated <- AddModuleScore(object = cd11b.integrated, features = list(gene_OA2), ctrl = 100, name = feature) 

feature="Aging_OA3"
cd11b.integrated <- AddModuleScore(object = cd11b.integrated, features = list(gene_OA3), ctrl = 100, name = feature) 
```

### Plot Violin plot for mouse microglia feature from multiple studies
```{r, fig.width=3.5, fig.height= 7}
meta <- cd11b.integrated@meta.data

meta_tidy <- meta %>% 
  select(Strain, Genotype, final_clusters:Aging_OA31) %>% 
  gather(key="study", value = "z_score", -Strain, -Genotype, -final_clusters)

study_ms <- c("Homeostatic_feature1", "DAM_feature1", "ARM_feature1", "Aging_OA21", "IRM_feature1", "Aging_OA31", "CPM_feature1")
study_ms_lab <- c("Homeostatic", "DAM_Amit", "ARM_de_Strooper", "AgingOA2_Stevens", "IRM_de_Strooper", "AgingOA3_Stevnes", "CPM_de_Strooper")

meta_tidy <- meta_tidy %>% 
  filter(study %in% study_ms) %>% 
  mutate(study = factor(study, levels = study_ms, labels = study_ms_lab))

meta_tidy%>% 
  ggplot(aes(y=z_score, x= final_clusters, color= final_clusters)) +
    facet_grid(study ~ ., scales= "fixed") +
    geom_violin() +
    geom_boxplot(width=0.15, outlier.shape = NA, color = "black", alpha=0.7) +
    theme_bw()+
    theme(legend.position = "none", 
          axis.title = element_blank(), 
          strip.text = element_text(face = "bold", size = 8, family = "Arial"),
          axis.text = element_text(size = 8))
ggsave(paste(output_path, "Vln_ms_multi_comp.png", sep = ""), dpi = 300, width = 3.5, height = 5)
```


## Comparing the z-score across clusters for each feature (two-way ANOVA)

### DAM_feature1 (Amit)
```{r }

clusters <- levels(meta$final_clusters) %>% as.list()

####single function for stain and GT interaction
data = meta %>% filter(final_clusters %in% clusters[[2]])
aov_object = aov(DAM_feature1 ~ Strain*Genotype , data=data)
aov.pvals = summary(aov_object)
aov.pvals= aov.pvals[[1]][5] %>% t() %>% as.data.frame()
names(aov.pvals) <- c("Strain", "Genotype", "Strain_Genotype", "Residuals")
aov.pvals <- aov.pvals %>% 
  select(-Residuals) %>% 
  mutate(Cluster = clusters[2] %>% as.character())

### wrap into function for DAM_feature1
aov_StrainGT <- function(cluster, data){
  data = data %>% filter(final_clusters %in% clusters)
  aov_object = aov(DAM_feature1 ~ Strain*Genotype, data=data)
  aov.pvals = summary(aov_object)
  aov.pvals= aov.pvals[[1]][5] %>% t() %>% as.data.frame()
  names(aov.pvals) <- c("Strain", "Genotype", "Strain_Genotype", "Residuals")
  aov.pvals <- aov.pvals %>% 
    select(-Residuals) %>% 
    mutate(Cluster = cluster %>% as.character())
  return(aov.pvals)
}

aov_StrainGT_object <- function(cluster, data){
  data = data %>% filter(final_clusters %in% clusters)
  aov_object = aov(DAM_feature1 ~ Strain*Genotype, data=data)
  return(aov_object)
}

aov_StrainGT_table <- clusters %>% map_df(aov_StrainGT, data=meta)
aov_StrainGT_table <- aov_StrainGT_table %>% mutate_if(is.double, p.adjust)

aov_StrainGT_table$Cluster[aov_StrainGT_table$Strain_Genotype<0.05]

aov_StrainGT_table$Cluster[aov_StrainGT_table$Strain<0.05]

aov_StrainGT_table$Cluster[aov_StrainGT_table$Genotype<0.05]

```

All clusters have strain, genotype and strain by genotype effects for DAM_feature1, we focus on cluster 6 and cluster 12. 

```{r}

aov_object_list <- clusters %>% map(aov_StrainGT_object, data=meta)
names(aov_object_list) <- clusters %>% unlist()

TukeyHSD(aov_object_list[["6"]]) %>% .$`Strain:Genotype` %>% data.frame(.,cluster="6")
TukeyHSD(aov_object_list[["12"]]) %>% .$`Strain:Genotype` %>% data.frame(.,cluster="12")

```


### IRM_feature1 (Bart de Strooper)

```{r }

clusters <- levels(meta$final_clusters) %>% as.list()
sig.level <- 1e-10

### function for IRM_feature1
aov_StrainGT <- function(cluster, data){
  data = data %>% filter(final_clusters %in% clusters)
  aov_object = aov(IRM_feature1 ~ Strain*Genotype, data=data) ## IRM
  aov.pvals = summary(aov_object)
  aov.pvals= aov.pvals[[1]][5] %>% t() %>% as.data.frame()
  names(aov.pvals) <- c("Strain", "Genotype", "Strain_Genotype", "Residuals")
  aov.pvals <- aov.pvals %>% 
    select(-Residuals) %>% 
    mutate(Cluster = cluster %>% as.character())
  return(aov.pvals)
}

aov_StrainGT_object <- function(cluster, data){
  data = data %>% filter(final_clusters %in% clusters)
  aov_object = aov(IRM_feature1 ~ Strain*Genotype, data=data)  ## IRM
  return(aov_object)
}

aov_StrainGT_table <- clusters %>% map_df(aov_StrainGT, data=meta)
aov_StrainGT_table <- aov_StrainGT_table %>% mutate_if(is.double, p.adjust)

aov_StrainGT_table$Cluster[aov_StrainGT_table$Strain_Genotype<sig.level]

aov_StrainGT_table$Cluster[aov_StrainGT_table$Strain<sig.level]

aov_StrainGT_table$Cluster[aov_StrainGT_table$Genotype<sig.level]

```


```{r}
aov_object_list <- clusters %>% map(aov_StrainGT_object, data=meta)
names(aov_object_list) <- clusters %>% unlist()

TukeyHSD(aov_object_list[["7"]]) %>% .$`Strain:Genotype` %>% data.frame(.,cluster="7")

```
