---
title: 'Comparing cluster marker genes across strains with reported microglia state: DAM (cluster
  6 and 12), IRM, and cluster10'
output: 
  html_document:
    keep_md: true
---

```{r global_options}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```


```{r}
library(tidyverse)
library(UpSetR)
library(Seurat)
library(cowplot)
library(Vennerable)
library(extrafont)
#font_import()

```

### set up marker gene sets for each cluster
```{r}

output_path <- "../03_results/07b_mic_compare_strain_ms/"
file_names <- c("DE_6_vs_H.txt", "DE_7_vs_H.txt", "DE_8_vs_H.txt", "DE_10_vs_H.txt", "DE_12_vs_H.txt")
file_dir <- file.path("../03_results/05b_makers_vs_Homeos/", file_names)
cluster_name <- c("DAM1", "IRM", "Cluster8", "Cluster10", "DAM2")

strain=c("B6", "CAST", "PWK", "WSB")

#load the DE gene data sets for all clusters
df <- file_dir %>% map(read_delim, delim = "\t")
names(df) <- cluster_name

## check the DAM1 DE genes from all strains (combined table from all strains)
for(i in seq_along(strain)){
  a <- df[["DAM1"]] %>% select(Symbol, contains(strain[i])) %>% 
    filter(str_detect(Symbol, "^Gm", negate = TRUE)) %>% 
    filter_at(vars(contains("p_val_adj")), any_vars(-log10(.)>25)) %>% 
    filter_at(vars(contains("logFC")), any_vars(.>0.25))
  print(dim(a))
}

```

### wrap into a function to retrieve DE gene sets 
```{r}

cluster_DE <- function(logFDR_cut, logFC_cut, df){
  # input:
  # logFDR_cut, threthold of -log10FDR
  # logFC_cut, threthold of log2FC
  # df: DE gene table for each cluster
  # output: 
  # a list containing marker genes for a given cluster for all strains 
  # strain: global environment
  DE_list <- vector(mode = "list", length = length(strain))
  
  if(logFC_cut>0){
    for(i in seq_along(strain)){
    DE_list[[i]] <- df %>% select(Symbol, contains(strain[i])) %>% 
      filter(str_detect(Symbol, "^Gm", negate = TRUE)) %>% 
      filter_at(vars(contains("p_val_adj")), any_vars(-log10(.)>logFDR_cut)) %>% 
      filter_at(vars(contains("logFC")), any_vars(.>logFC_cut)) %>% 
      select(Symbol) %>% unlist()
    }
  }else{
    for(i in seq_along(strain)){
    DE_list[[i]] <- df %>% select(Symbol, contains(strain[i])) %>% 
      filter(str_detect(Symbol, "^Gm", negate = TRUE)) %>% 
      filter_at(vars(contains("p_val_adj")), any_vars(-log10(.)>logFDR_cut)) %>% 
      filter_at(vars(contains("logFC")), any_vars(.<logFC_cut)) %>% 
      select(Symbol) %>% unlist()
    }
  }
  
  names(DE_list) <- strain
  return(DE_list)
}


```




```{r}
# to retrieve marker gene for each cluster across strains, can use the following expression.
DE_DAM6_list <- cluster_DE(18, 0.25, df[["DAM1"]])
    DE_DAM6_list %>% map_df(length)
    
DE_IRM_list <- cluster_DE(8, 0.25, df[["IRM"]])
    DE_IRM_list %>% map_df(length)
    
DE_DAM12_list <- cluster_DE(18, 0.25, df[["DAM2"]])
    DE_DAM12_list %>% map_df(length)
    
DE_cl10_list <- cluster_DE(8, 0.25, df[["Cluster10"]])
    DE_cl10_list  %>% map_df(length)

```



### Marker gene sets from other studies
##### set up marker gene sets for comparison (DAM: Amit, ARM: Strooper, IRM: Strooper, Aging_OA2: Steven, Aging_OA3: Steven)
1. Amits DAM genes (and Homeostatic microglia genes: negative DAM markers)
```{r}

sig_levels <- c(18, 0.25)   
# logFDR_cut, threthold of -log10FDR # logFC_cut, threthold of log2FC

df_comp <- readxl::read_xlsx("../02_data/mic_comp_Amit/Amit_HvsDAM.xlsx", col_types=c("text", rep("numeric",4)))
names(df_comp) <- c("Symbol", "UMI_H", "log2FC", "log10p", "log10FDR")
df_comp <- mutate(df_comp, 
                      p_val=10^(-log10p), 
                      p_val_adj=10^(-log10FDR))
    ## DAM markers of all strains
df_DAM <- df_comp %>% 
  filter(log10FDR>sig_levels[1], log2FC>sig_levels[2], str_detect(Symbol, "^Gm", negate = TRUE)) %>% 
  arrange(desc(log10FDR)) ## top 58 genes
df_DAM
gene_DAM <- df_DAM$Symbol %>% unlist()


```

2. Strooper ARM genes (and Homeostatic microglia genes: negative ARM markers)
```{r}
#ARM
sig_levels <- c(6, 0.25)
df_comp <- read_delim("../03_results/06a_mic_Strooper/Seurat_Strooper_markers.txt", delim = "\t")
df_comp <- df_comp %>% 
  filter(cluster==2) %>%  # ARM cluster 2
  filter(-log10(p_val_adj)>sig_levels[1], avg_logFC>sig_levels[2], str_detect(symbol, "^Gm", negate = TRUE)) %>% 
  arrange(desc(-log10(p_val_adj)))## top 59 genes
gene_ARM <- df_comp$gene %>% unlist()
df_comp


```


3. Strooper IRM genes
```{r}
sig_levels <- c(6, 0.25)
df_comp <- read_delim("../03_results/06a_mic_Strooper/Seurat_Strooper_markers.txt", delim = "\t")
df_comp <- df_comp %>% 
  filter(cluster==7) %>%  # IRM cluster 7
  filter(-log10(p_val_adj)>sig_levels[1], avg_logFC>sig_levels[2], str_detect(symbol, "^Gm", negate = TRUE)) %>% 
  arrange(desc(-log10(p_val_adj)))## top 53 genes
gene_IRM <- df_comp$gene %>% unlist()
df_comp

```

4. Stevens Ag_OA2 (aging related)
```{r}
# data loading copied from "mic_comp_addmodule"
input_file <- "../02_data/mic_comp_Stevens/Stevens_cluster_marker.xlsx"
M_stevens_sheet_name <- readxl::excel_sheets(input_file)
sheet <- 1:(length(M_stevens_sheet_name)-1) # do not include the last Transcription factors
df_comp <- vector(mode = "list", length = length(sheet))

for(i in sheet){
  df_comp[[i]] <- readxl::read_xlsx(input_file, sheet = i, skip = 2)
  df_comp[[i]] <- df_comp[[i]] %>% 
    select(1:6)%>% 
    mutate(cluster=M_stevens_sheet_name[i]) 
}

df_comp <- do.call(rbind, df_comp)

colnames(df_comp) <- c("gene", "FC", "p_val", "p_val_adj", "pct_1", "pct_2", "cluster")

df_comp <- df_comp %>% mutate(cluster=str_remove_all(cluster, "Cluster|\\(|\\)| "), 
                              p_val=as.double(p_val),
                              p_val_adj=as.double(p_val_adj))

sig_levels <- c(18, 0.25)
gene_OA2 <- df_comp %>% 
    filter(cluster=="AgingOA2", -log10(p_val_adj)>sig_levels[1], log2(FC)>sig_levels[2], str_detect(gene, "^Gm", negate = TRUE)) %>% 
    arrange(p_val_adj, desc(FC)) %>% head(50) %>% .$gene %>% unlist() 

gene_OA2

```

5. Stevens Ag_OA3 (IRM similar)
```{r}
sig_levels <- c(0.05, 0.25)
gene_OA3 <- df_comp %>% 
    filter(cluster=="AgingOA3", -log10(p_val_adj)>sig_levels[1], log2(FC)>sig_levels[2], str_detect(gene, "^Gm", negate = TRUE)) %>% 
    arrange(p_val_adj, desc(FC)) %>% .$gene %>% unlist() # 31

gene_OA3

```


## Upset plot 1: DAM6 overlapping with Amit DAM & Strooper ARM
```{r}

comp <- "DAM6_Amit_Strooper"
DE_cluster_list <- DE_DAM6_list

DE_cluster_list[["Amit"]] <- gene_DAM
DE_cluster_list[["de Strooper"]] <- gene_ARM 

file_name <- paste(output_path, "upset_", comp, ".png", sep = "")

x <- upset(fromList(DE_cluster_list), 
      nsets = length(DE_cluster_list), 
      sets = names(DE_cluster_list),
      queries = list(list(query = intersects, params = list(names(DE_cluster_list)), color = "orange", active = TRUE),
                     list(query = intersects, params = list(names(DE_cluster_list)[-5:-6]), color = "grey", active = TRUE)),
      order.by = "freq",
      keep.order = TRUE,
      sets.bar.color = "blue", text.scale=1.5)
x

png(filename = file_name, width = 8, height = 4.5, units = "in", res = 300)
x
dev.off()

```

### Export intersecting DAM6 gene list: 
```{r}
### Upset plot 1: DAM6 overlapping with Amit DAM & Strooper ARM

## Because this upset has 6 lists, so venn() function in gplots package doesn't work. use Vennerable for up to 9 sets

#library(devtools)
#install_github("js229/Vennerable")
library(Vennerable)


tmp <- Venn(DE_cluster_list)
# view elements:
tmp@IntersectionSets$`111111` ## all overlap
tmp@IntersectionSets$`111100` ## B6,CAST,PWK,WSB overlap, but not in Strooper/Amit
tmp@IntersectionSets$`010011` ## Strooper&Amit, & CAST 
tmp@IntersectionSets$`110011` ## Strooper&Amit, & CAST&B6
tmp@IntersectionSets$`100011` ## Strooper&Amit, & B6
tmp@IntersectionSets$`100111` ## Strooper&Amit, & B6 & WSB
tmp@IntersectionSets$`110111` ## Strooper&Amit, & B6 & WSB & CAST
tmp@IntersectionSets$`101111` ## Strooper&Amit, & B6 & WSB & PWK
tmp@IntersectionSets$`101011` ## Strooper&Amit, & B6 & PWK
tmp@IntersectionSets$`111011` ## Strooper&Amit, & B6 & PWK & CAST

# save core DAM6 genes for score evaluation 
core_DAM6 <- tmp@IntersectionSets$`111111` 
# upset_DAM1 <- tmp
#write_delim(DE_cluster_intersect, paste(output_path, "upset_", comp, "_intersect.txt", sep = ""), delim = "\t")


```


### load seurat object to add module scores based on core genes of DAM 6 
```{r}
cd11b.integrated <- readRDS("../02_data/intermediate_rds/mg_int.rds") 

## combine homeostatic microglia
cd11b.integrated$final_clusters <-  ifelse(cd11b.integrated$seurat_clusters %in% 0:5,"H",
                                           cd11b.integrated$seurat_clusters %>% as.character())
cd11b.integrated$final_clusters <- factor(cd11b.integrated$final_clusters, 
                                          levels = c("H", "6", "7", "8", "9", "10", "11", "12"))
cd11b.integrated$Genotype <- factor(cd11b.integrated$Genotype, levels = c("WT", "APP/PS1"))

cd11b.integrated$Strain <- gsub("B6J", "B6", cd11b.integrated$Strain)
cd11b.integrated$Strain <- factor(cd11b.integrated$Strain, levels = c("B6", "CAST", "PWK", "WSB"))

cd11b.integrated$Group <- gsub("B6J", "B6", cd11b.integrated$Group)

cd11b.integrated$Group <- factor(cd11b.integrated$Group, 
                                 levels=c("B6_WT", "B6_APP/PS1", "CAST_WT", "CAST_APP/PS1", "PWK_WT","PWK_APP/PS1", "WSB_WT", "WSB_APP/PS1"))

```


### add module score for all core clusters (DAM/ARM, IRM/Aging_OA3, Aging_OA2)
```{r}
cd11b.integrated <- AddModuleScore(object = cd11b.integrated, features = list(core_DAM6), ctrl = 100, name = 'core_DAM6') 


```


#### plot
```{r, fig.width=2.5, fig.height=5}
meta <- cd11b.integrated@meta.data

clusters <- levels(meta$final_clusters) %>% as.list()

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") # color blind friendly Palette (grey)

```

### List examples of marker genes of all kinds splits by strain and genotype

### DAM 
```{r, fig.width=9.6, fig.height=3.3}

genes <- c("Cst7", "Clec7a", "Cd9", "Lpl", "Apoe","Cadm1", "Ctsa", "Trem2", "Cd68", "Ctsl", "Capg", "Serpine2")

library(extrafont)
#font_import()

Idents(cd11b.integrated) <- "Strain"
p <- genes %>% 
  map(~VlnPlot(cd11b.integrated %>% subset(subset=final_clusters %in% c("6")), features = ., pt.size = 0, split.by = "Genotype", ncol = 4, cols =  c("#E69F00", "#999999")) + 
        theme(legend.position = "none", 
              axis.title = element_blank(), 
              axis.text.x = element_text(size = 8),
              axis.text.y = element_text(size = 10), 
              title = element_text(size = 10, family = "Arial")
              )) 

plot_grid(p[[1]], p[[2]], p[[3]], p[[4]], p[[5]], p[[6]], p[[7]], p[[8]], p[[9]], p[[10]], p[[11]], p[[12]], align = c("hv"), nrow =2, ncol = 6)

ggsave(paste(output_path, "Vln_gene_", comp, ".png", sep = ""), dpi = 300, width = 9.6, height = 3.3)

# just plot a gene figure with the legend
x <- VlnPlot(cd11b.integrated %>% subset(subset=final_clusters %in% c("6")), features = "Lpl", pt.size = 0, split.by = "Genotype", ncol = 4, cols =  c("#E69F00", "#999999"))

ggsave(paste(output_path, "Vln_lpl_", comp, ".png", sep = ""), x, dpi = 300, width = 8, height = 3.3)

```


## Upset plot 1.1 : Cluster12 overlapping with Amit and de Strooper
```{r, fig.width=8, fig.height=4.5}
comp <- "DAM12_Amit_Strooper"
DE_cluster_list <- DE_DAM12_list

DE_cluster_list[["Amit"]] <- gene_DAM
DE_cluster_list[["de Strooper"]] <- gene_ARM 

file_name <- paste(output_path, "upset_", comp, ".png", sep = "")

x <- upset(fromList(DE_cluster_list), 
      nsets = length(DE_cluster_list), 
      sets = names(DE_cluster_list),
      queries = list(list(query = intersects, params = list(names(DE_cluster_list)), color = "orange", active = TRUE),
                     list(query = intersects, params = list(names(DE_cluster_list)[-5:-6]), color = "grey", active = TRUE)),
      order.by = "freq",
      keep.order = TRUE,
      sets.bar.color = "blue", text.scale=1.5)
x

png(filename = file_name, width = 8, height = 4.5, units = "in", res = 300)
x
dev.off()

```

### Export intersecting DAM12 gene list: 
```{r}
library(Vennerable)

tmp <- Venn(DE_cluster_list)
# view elements:
tmp@IntersectionSets$`111111` ## all overlap
tmp@IntersectionSets$`111100` ## B6,CAST,PWK,WSB overlap, but not in de Strooper/Amit
tmp@IntersectionSets$`000011` ## Not in B6,CAST,PWK,WSB overlap, but in de Strooper/Amit
tmp@IntersectionSets$`110111` ## Strooper&Amit, & B6 & WSB & CAST
tmp@IntersectionSets$`110011` ## Strooper&Amit, & CAST & B6

tmp@IntersectionSets$`001011` ## Strooper&Amit, & PWK
tmp@IntersectionSets$`010011` ## Strooper&Amit, & CAST
tmp@IntersectionSets$`110011` ## Strooper&Amit, & CAST&B6
tmp@IntersectionSets$`011111` ## Strooper&Amit, & CAST & PWK & WSB
tmp@IntersectionSets$`101111` ## Strooper&Amit, & B6 & PWK & WSB
tmp@IntersectionSets$`111011` ## Strooper&Amit, & B6 & PWK & CAST


# save core DAM6 genes for score evaluation 
core_DAM12 <- tmp@IntersectionSets$`111111` 
# upset_DAM1 <- tmp
#write_delim(DE_cluster_intersect, paste(output_path, "upset_", comp, "_intersect.txt", sep = ""), delim = "\t")
```
### add module score for all core genes of cluster 12
```{r}
cd11b.integrated <- AddModuleScore(object = cd11b.integrated, features = list(core_DAM12), ctrl = 100, name = 'core_DAM12') 


```



### List examples of marker genes of all kinds splits by strain and genotype
### DAM12
```{r, fig.width=9.6, fig.height=3.3}

genes <- c("Tyrobp", "Cst7", "Clec7a", "Ccl6", "B2m", "Ctsd", "Ccl3", "Lpl", "Trem2", "Axl", "Cd9", "Apoe")

library(extrafont)
#font_import()

Idents(cd11b.integrated) <- "Strain"
p <- genes %>% 
  map(~VlnPlot(cd11b.integrated %>% subset(subset=final_clusters %in% c("12")), features = ., pt.size = 0, split.by = "Genotype", ncol = 4, cols =  c("#E69F00", "#999999")) + 
        theme(legend.position = "none", 
              axis.title = element_blank(), 
              axis.text.x = element_text(size = 8),
              axis.text.y = element_text(size = 10), 
              title = element_text(size = 10, family = "Arial")
              )) 

plot_grid(p[[1]], p[[2]], p[[3]], p[[4]], p[[5]], p[[6]], p[[7]], p[[8]], p[[9]], p[[10]], p[[11]], p[[12]], align = c("hv"), nrow =2, ncol = 6)

ggsave(paste(output_path, "Vln_gene_", comp, ".png", sep = ""), dpi = 300, width = 9.6, height = 3.3)

```



## Upset plot 2: IRM7 overlapping with Strooper IRM & Stevens Aging_OA3
```{r}
comp <- "IRM7_Strooper_Stevens"
DE_cluster_list <- DE_IRM_list

DE_cluster_list[["de Strooper"]] <- gene_IRM
DE_cluster_list[["Stevens"]] <- gene_OA3

file_name <- paste(output_path, "upset_", comp, ".png", sep = "")

x <- upset(fromList(DE_cluster_list), 
      nsets = length(DE_cluster_list), 
      sets = names(DE_cluster_list),
      queries = list(list(query = intersects, params = list(names(DE_cluster_list)), color = "orange", active = TRUE)),
      order.by = "freq",
      keep.order = TRUE,
      sets.bar.color = "blue", text.scale=1.5)
x

png(filename = file_name, width = 8, height = 4.5, units = "in", res = 300)
x
dev.off()


```

### Export intersecting IRM7 gene list: 
```{r}

tmp <- Venn(DE_cluster_list)
# view elements:
tmp@IntersectionSets$`111111` ## all overlap
tmp@IntersectionSets$`111110` ## B6,CAST,PWK,WSB overlap with Strooper but not Stevens
tmp@IntersectionSets$`111101` ## B6,CAST,PWK,WSB overlap with Strooper but not Stevens

tmp@IntersectionSets$`000111` ## Strooper/Stevens marker, WSB only
tmp@IntersectionSets$`100111` ## Strooper/Stevens marker, WSB & B6
tmp@IntersectionSets$`011011` ## Strooper/Stevens marker, PWK & CAST
tmp@IntersectionSets$`101111` ## Strooper/Stevens marker, WSB & B6 & PWK, not CAST
tmp@IntersectionSets$`111011` ## Strooper/Stevens marker, CAST & B6 & PWK, not WSB
tmp@IntersectionSets$`110111` ## Strooper/Stevens marker, WSB & B6 & CAST, not PWK

# save core IRM7 genes for score evaluation 
core_IRM7 <- tmp@IntersectionSets$`111111` 

```

### add module score for core markers IRM/Aging_OA3
```{r}
cd11b.integrated <- AddModuleScore(object = cd11b.integrated, features = list(core_IRM7), ctrl = 100, name = 'core_IRM7') 
```


### List examples of marker genes of all kinds splits by strain and genotype

### IRM 
```{r, fig.width=9.6, fig.height=3.3}

genes <- c("Rtp4", "Ifit3", "Ifitm3", "Irf7", "Bst2", "Oasl2", "Trim30a", "Stat1", "Oas1a", "Ccl12", "Tor3a", "Ifi204")

#library(extrafont)
#font_import()

Idents(cd11b.integrated) <- "Strain"
p <- genes %>% 
  map(~VlnPlot(cd11b.integrated %>% subset(subset=final_clusters %in% c("7")), features = ., pt.size = 0, split.by = "Genotype", ncol = 4, cols =  c("#E69F00", "#999999")) + 
        theme(legend.position = "none", 
              axis.title = element_blank(), 
              axis.text.x = element_text(size = 8),
              axis.text.y = element_text(size = 10), 
              title = element_text(size = 10, family = "Arial")
              )) 

plot_grid(p[[1]], p[[2]], p[[3]], p[[4]], p[[5]], p[[6]], p[[7]], p[[8]], p[[9]], p[[10]], p[[11]], p[[12]], align = ("hv"), nrow =2, ncol = 6)

ggsave(paste(output_path, "Vln_gene_", comp, ".png", sep = ""), dpi = 300, width = 9.6, height = 3.3)

```


## Upset plot 3: cluster10 overlapping with Stevens Aging_OA2
```{r}
comp <- "cl10_Stevens_OA2"
DE_cluster_list <- DE_cl10_list

DE_cluster_list[["Stevens"]] <- gene_OA2

file_name <- paste(output_path, "upset_", comp, ".png", sep = "")

x <- upset(fromList(DE_cluster_list), 
      nsets = length(DE_cluster_list), 
      sets = names(DE_cluster_list),
      queries = list(list(query = intersects, params = list(names(DE_cluster_list)), color = "orange", active = TRUE)),
      order.by = "freq",
      keep.order = TRUE,
      sets.bar.color = "blue", text.scale=1.5)
x

png(filename = file_name, width = 8, height = 4.5, units = "in", res = 300)
x
dev.off()


```

### Export intersecting cluster10 gene list: 
```{r}

tmp <- Venn(DE_cluster_list)
# view elements:
tmp@IntersectionSets$`11111` ## all overlap
tmp@IntersectionSets$`11110` ## B6,CAST,PWK,WSB overlap, but not in Stevens

tmp@IntersectionSets$`00001` ## Not in B6,CAST,PWK,WSB overlap, but in Stevens

tmp@IntersectionSets$`10001` ## Stevens, & B6
tmp@IntersectionSets$`11101` ## Stevens, & B6 & PWK & CAST

tmp@IntersectionSets$`00101` ## Stevens, & PWK

tmp@IntersectionSets$`01001` ## Stevens, & CAST
tmp@IntersectionSets$`01101` ## Stevens, & CAST & PWK

tmp@IntersectionSets$`01011` ## Stevens, & CAST & WSB
tmp@IntersectionSets$`10111` ## Stevens, & B6 & PWK & WSB


# save core DAM6 genes for score evaluation 
core_cluster10 <- tmp@IntersectionSets$`11111` 

```

### add module score for core markers cluster 10/ Aging_OA2
```{r}
cd11b.integrated <- AddModuleScore(object = cd11b.integrated, features = list(core_cluster10), ctrl = 100, name = 'core_cluster10') 
```



### Cluster 10 gene violin plots in each strain and genotype
```{r, fig.width=9.6, fig.height=3.3}

genes <- c("Ccl3", "Ccl4", "Cd83", "C3ar1", "Atf3", "Lpl", "Lgals3", "Cd68", "Tlr2", "C5ar1", "Prdx1", "Lyz2")

#library(extrafont)
#font_import()

Idents(cd11b.integrated) <- "Strain"
p <- genes %>% 
  map(~VlnPlot(cd11b.integrated %>% subset(subset=final_clusters %in% c("10")), features = ., pt.size = 0, split.by = "Genotype", ncol = 4, cols =  c("#E69F00", "#999999")) + 
        theme(legend.position = "none", 
              axis.title = element_blank(), 
              axis.text.x = element_text(size = 8),
              axis.text.y = element_text(size = 10), 
              title = element_text(size = 10, family = "Arial")
              )) 

plot_grid(p[[1]], p[[2]], p[[3]], p[[4]], p[[5]], p[[6]], p[[7]], p[[8]], p[[9]], p[[10]], p[[11]], p[[12]], align = ("hv"), nrow =2, ncol = 6)

ggsave(paste(output_path, "Vln_gene_", comp, ".png", sep = ""), dpi = 300, width = 9.6, height = 3.3)

```

#### Homeostatic microglia expression

Homoestatic marker paper: 
Amit, Stooper, Stevens.
Christoph glass: https://science.sciencemag.org/content/early/2017/05/24/science.aal3222?versioned=true
Oleg : https://www.nature.com/articles/s41583-018-0057-5

```{r}
gene_homeo <- c("Tmem119", "Cx3cr1", "P2ry12", "P2ry13", "Hexb", "Olfml3",  "Selplg", "Siglech", "Csf1r", "Cst3", "Sparc", "C1qa", "C1qb", "C1qc", "Tmsb4x", "Sall1", "Fcrls", "Gpr34", "Spi1",  "Mafb", "Maf", "Mef2a", "Irf8")

```

```{r}
cd11b.integrated <- AddModuleScore(object = cd11b.integrated, features = list(gene_homeo), ctrl = 100, name = 'homeo') 
```

```{r , fig.width= 3, fig.height=6}
meta <- cd11b.integrated@meta.data
saveRDS(meta, paste(output_path, "meta_mouse_strain_score.rds", sep = ""))

```


### homeostatic microglia markers
```{r, fig.width=6.4, fig.height=5}

genes <- c("Tmem119", "Cx3cr1", "P2ry12", "Sall1", "Gpr34", "Spi1",  "Mafb", "Maf", "Mef2a", "Irf8",  "Fcrls", "Olfml3")


Idents(cd11b.integrated) <- "Strain"
p <- genes %>% 
  map(~VlnPlot(cd11b.integrated %>% subset(subset=final_clusters %in% c("H")), features = ., pt.size = 0, split.by = "Genotype", ncol = 4, cols =  c("#E69F00", "#999999")) + 
        theme(legend.position = "none", 
              axis.title = element_blank(), 
              axis.text.x = element_text(size = 8),
              axis.text.y = element_text(size = 10), 
              title = element_text(size = 10, family = "Arial")
              )) 

plot_grid(p[[1]], p[[2]], p[[3]], p[[4]], p[[5]], p[[6]], p[[7]], p[[8]], p[[9]], p[[10]], 
          p[[11]], p[[12]],
          align = ("hv"), nrow =3, ncol = 4)
# width = 8, height = 3.3 2x5

ggsave(paste(output_path, "Vln_gene_", comp, ".png", sep = ""), dpi = 300, width = 6.4, height = 5)

```

## z-score plotting using split violin and grouped by cluster
```{r ,fig.width= 3, fig.height=6}
# can run this chunck without runing previous chuncks
library(tidyverse)

meta <- readRDS(paste(output_path, "meta_mouse_strain_score.rds", sep = ""))
source("~/code/wild_AD_mic_scRNA/05_function/split_ViolinBox_plot.R")

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") # color blind friendly Palette (grey)

## plot Homoestatic microglia z-score
comp <- "homeo1"
cluster <- c("H", "8", "6")
strain <- c("B6", "CAST", "PWK", "WSB") %>% rev()
width_height <- c(3, 5)

meta %>%
  filter(final_clusters %in% cluster) %>% 
  mutate(final_clusters = factor(final_clusters, levels = cluster),
         Strain = factor(Strain, levels = strain)) %>% 
  ggplot(aes(y=homeo1, x= Strain, fill= Genotype)) +   # change y to the same as comp but no quotation mark
    facet_grid(final_clusters ~ ., scales= "fixed") +
    geom_split_violin(trim = TRUE) +
    geom_boxplot(width = 0.25, notch = FALSE, notchwidth = .4, outlier.shape = NA, coef=0)+
    scale_fill_manual(values=cbPalette) + 
    coord_flip() +
    theme_bw()+
    labs(x=NULL, y="z-score") +
    theme(legend.position = "bottom", 
          strip.text = element_text(face = "bold", size = 12, family = "Arial"),
          axis.text = element_text(size = 10), 
          axis.text.y = element_text(face = "bold"),
          axis.title.x = element_text(face = "bold", size = 12))

ggsave(paste(output_path, "Vln2_", comp, ".png", sep = ""), dpi = 300, width = width_height[1], height = width_height[2])


## plot DAM6 features
comp <- "core_DAM61"
cluster <- c("H", "6")
strain <- c("B6", "CAST", "PWK", "WSB") %>% rev()
width_height <- c(2.5, 5)

meta %>%
  filter(final_clusters %in% cluster) %>% 
  mutate(final_clusters = factor(final_clusters, levels = cluster),
         Strain = factor(Strain, levels = strain)) %>% 
  ggplot(aes(y=core_DAM61, x= Strain, fill= Genotype)) +   # change y to the same as comp but no quotation mark
    facet_grid(final_clusters ~ ., scales= "fixed") +
    geom_split_violin(trim = TRUE) +
    geom_boxplot(width = 0.25, notch = FALSE, notchwidth = .4, outlier.shape = NA, coef=0)+
    scale_fill_manual(values=cbPalette) + 
    coord_flip() +
    theme_bw()+
    labs(x=NULL, y="z-score") +
    theme(legend.position = "bottom", 
          strip.text = element_text(face = "bold", size = 12, family = "Arial"),
          axis.text = element_text(size = 10), 
          axis.text.y = element_text(face = "bold"),
          axis.title.x = element_text(face = "bold", size = 12))

ggsave(paste(output_path, "Vln2_", comp, ".png", sep = ""), dpi = 300, width = width_height[1], height = width_height[2])

## plot IRM7 features
comp <- "core_IRM71"
cluster <- c("H", "7")
strain <- c("B6", "CAST", "PWK", "WSB") %>% rev()
width_height <- c(2.5, 5)

meta %>%
  filter(final_clusters %in% cluster) %>% 
  mutate(final_clusters = factor(final_clusters, levels = cluster),
         Strain = factor(Strain, levels = strain)) %>% 
  ggplot(aes(y=core_IRM71, x= Strain, fill= Genotype)) +   # change y to the same as comp but no quotation mark
    facet_grid(final_clusters ~ ., scales= "fixed") +
    geom_split_violin(trim = TRUE) +
    geom_boxplot(width = 0.25, notch = FALSE, notchwidth = .4, outlier.shape = NA, coef=0)+
    scale_fill_manual(values=cbPalette) + 
    coord_flip() +
    theme_bw()+
    labs(x=NULL, y="z-score") +
    theme(legend.position = "bottom", 
          strip.text = element_text(face = "bold", size = 12, family = "Arial"),
          axis.text = element_text(size = 10), 
          axis.text.y = element_text(face = "bold"),
          axis.title.x = element_text(face = "bold", size = 12))

ggsave(paste(output_path, "Vln2_", comp, ".png", sep = ""), dpi = 300, width = width_height[1], height = width_height[2])


## plot DAM12 features
comp <- "core_DAM121"
cluster <- c("H", "12")
strain <- c("B6", "CAST", "PWK", "WSB") %>% rev()
width_height <- c(2.5, 5)

meta %>%
  filter(final_clusters %in% cluster) %>% 
  mutate(final_clusters = factor(final_clusters, levels = cluster),
         Strain = factor(Strain, levels = strain)) %>% 
  ggplot(aes(y=core_DAM121, x= Strain, fill= Genotype)) +   # change y to the same as comp but no quotation mark
    facet_grid(final_clusters ~ ., scales= "fixed") +
    geom_split_violin(trim = TRUE) +
    geom_boxplot(width = 0.25, notch = FALSE, notchwidth = .4, outlier.shape = NA, coef=0)+
    scale_fill_manual(values=cbPalette) + 
    coord_flip() +
    theme_bw()+
    labs(x=NULL, y="z-score") +
    theme(legend.position = "bottom", 
          strip.text = element_text(face = "bold", size = 12, family = "Arial"),
          axis.text = element_text(size = 10), 
          axis.text.y = element_text(face = "bold"),
          axis.title.x = element_text(face = "bold", size = 12))

ggsave(paste(output_path, "Vln2_", comp, ".png", sep = ""), dpi = 300, width = width_height[1], height = width_height[2])


## plot cluster 10 features
comp <- "core_cluster101"
cluster <- c("H", "10")
strain <- c("B6", "CAST", "PWK", "WSB") %>% rev()
width_height <- c(2.5, 5)

meta %>%
  filter(final_clusters %in% cluster) %>% 
  mutate(final_clusters = factor(final_clusters, levels = cluster),
         Strain = factor(Strain, levels = strain)) %>% 
  ggplot(aes(y=core_cluster101, x= Strain, fill= Genotype)) +   # change y to the same as comp but no quotation mark
    facet_grid(final_clusters ~ ., scales= "fixed") +
    geom_split_violin(trim = TRUE) +
    geom_boxplot(width = 0.25, notch = FALSE, notchwidth = .4, outlier.shape = NA, coef=0)+
    scale_fill_manual(values=cbPalette) + 
    coord_flip() +
    theme_bw() +
    labs(x=NULL, y="z-score") +
    theme(legend.position = "bottom", 
          strip.text = element_text(face = "bold", size = 12, family = "Arial"),
          axis.text = element_text(size = 10), 
          axis.text.y = element_text(face = "bold"),
          axis.title.x = element_text(face = "bold", size = 12))

ggsave(paste(output_path, "Vln2_", comp, ".png", sep = ""), dpi = 300, width = width_height[1], height = width_height[2])

```

### convert overlapping genes into gene list
```{r}

DE_cluster_ls_ls <- vector(mode = "list") # make a big list to store each intersection list

# DAM6
comp <- "DAM6_Amit_Strooper"
DE_cluster_list <- DE_DAM6_list
DE_cluster_list[["Amit"]] <- gene_DAM
DE_cluster_list[["de_Strooper"]] <- gene_ARM 
DE_cluster_ls_ls[[comp]] <- DE_cluster_list

# DAM12
comp <- "DAM12_Amit_Strooper"
DE_cluster_list <- DE_DAM12_list
DE_cluster_list[["Amit"]] <- gene_DAM
DE_cluster_list[["de_Strooper"]] <- gene_ARM 
DE_cluster_ls_ls[[comp]] <- DE_cluster_list

# IRM7
comp <- "IRM7_Strooper_Stevens"
DE_cluster_list <- DE_IRM_list
DE_cluster_list[["de_Strooper"]] <- gene_IRM
DE_cluster_list[["Stevens"]] <- gene_OA3
DE_cluster_ls_ls[[comp]] <- DE_cluster_list

#cluster10
comp <- "cl10_Stevens_OA2"
DE_cluster_list <- DE_cl10_list
DE_cluster_list[["Stevens"]] <- gene_OA2
DE_cluster_ls_ls[[comp]] <- DE_cluster_list

names(DE_cluster_ls_ls)

source("~/code/wild_AD_mic_scRNA/05_function/venn_intersects_upgrade.R")

intersect_table_list <- map(DE_cluster_ls_ls, venn_intersects_upgrade)

for (i in seq_along(intersect_table_list)){
  intersect_table_list[[i]] <- intersect_table_list[[i]] %>% 
    mutate(Orig_Symbol=as.character(Orig_Symbol),
           Intersections=paste("Intersection", as.character(Intersections), sep="_"))
}

walk2(intersect_table_list, paste(output_path, "Interset_", names(DE_cluster_ls_ls), ".txt", sep = ""), write_delim, delim="\t")

```
